---


---
```{r}
library(data.table)
library(tidyverse)
library(glue)

library(IRanges)
```

```{r}
work_dir <- '/project2/haky/temi/projects/TFXcan-snakemake'
sumstats_dir <- "/project2/haky/temi/projects/l-TFPred/experiments/AR_phenotypes/data/gwas_cat_sumstats"
```

```{r}
meta <- data.table::fread(glue('{sumstats_dir}/metadata.txt'))
meta
```

```{r}
# testosterone_gwas <- data.table::fread(file.path(gwas_folder, 'testosterone_gwas_sumstats.gcg.txt.gz'))
# #asthma_gwas <- data.table::fread(file.path(gwas_folder, 'asthma_gwas_sumstats.gcg.txt.gz'))
# prostate_cancer_gwas <- data.table::fread(file.path(gwas_folder, 'prostate_cancer_risk_gwas_sumstats.gcg.txt.gz')) 
# #psoriasis_gwas <- data.table::fread(file.path(gwas_folder, 'psoriasis_gwas_sumstats.txt.gz')) 
# bph_gwas <- data.table::fread(file.path(gwas_folder, 'bph_gwas_sumstats.gcg.txt.gz'))

# gwas_phenotypes_list <- list(testosterone = testosterone_gwas, prostate_cancer = prostate_cancer_gwas, bph = bph_gwas)
# lapply(gwas_phenotypes_list, colnames)
```

```{r}
gwas_folder <- '/project2/haky/temi/projects/l-TFPred/experiments/AR_phenotypes/data/processed_sumstats/'
out_folder <- '/project2/haky/temi/projects/TFPred/data/GWAS_sumstats'
gwas_folder <- '/project2/haky/temi/projects/l-TFPred/experiments/AR_phenotypes/data/processed_sumstats/'
run_script <- '/project2/haky/temi/projects/TFPred/scripts/clump_gwas_summary_statistics.sbatch'


out <- lapply(seq_along(meta$phenotype), function(i){
    pheno_folder <- glue('{out_folder}/{meta$phenotype[i]}')
    if(!dir.exists(pheno_folder)){
        dir.create(pheno_folder)
    }

    if(!file.exists(file.path(pheno_folder, glue('{meta$phenotype[i]}.gwas_sumstats.processed.txt')))){
        data.table::fread(file.path(gwas_folder, glue('{meta$phenotype[i]}_gwas_sumstats.gcg.txt.gz'))) %>%
            dplyr::mutate(SNP=paste(chrom, pos, A2, A1, sep='_'), P=p_value) %>%
            data.table::fwrite(., file.path(pheno_folder, glue('{meta$phenotype[i]}.gwas_sumstats.processed.txt')), col.names=T, row.names=F, sep='\t', quote=F)
    }
    
    print(glue('sbatch --export=phenotype={meta$phenotype[i]},pheno_folder={pheno_folder} {run_script}'))
        #dplyr::mutate(p_value_adj = p.adjust(p_value, method = 'bonferroni')) 

    # if(nrow(dt %>% dplyr::filter(p_value_adj < 0.01)) <= 10){
    #     return(NULL)
    # } else {
    #     dt %>% 
    #         dplyr::filter(p_value_adj < 0.01) %>% 
    #         data.table::fwrite(., file.path(work_dir, 'data/sumstats', glue('{meta$phenotype[i]}.gwas_sumstats.txt.gz')), col.names=T, row.names=F, quote=F, compress='gzip')

    #     dt %>% 
    #         dplyr::slice_max(order_by=desc(p_value_adj), n=1000) %>%
    #         data.table::fwrite(., file.path(work_dir, 'data/sumstats', glue('{meta$phenotype[i]}_no-hits.gwas_sumstats.txt.gz')), col.names=T, row.names=F, quote=F, compress='gzip')
        
    #     phenotypes <- c(meta$phenotype[i], 
    #         glue('{meta$phenotype[i]}_no-hits'))
    #     sumstats <- c(sumstat=glue('{meta$phenotype[i]}.gwas_sumstats.txt.gz'), 
    #         glue('{meta$phenotype[i]}_no-hits.gwas_sumstats.txt.gz'))
    # }

    # return(cbind(phenotype = phenotypes, sumstat=sumstats))
})
```

```{r}
print(glue('{run_script} '))
```


### Asthma

```{r}
asthma_folder <- '/project2/haky/temi/projects/TFPred/data/asthma'
asthma_phenotypes <- c('asthma_adults', 'asthma_children')

#gwas_folder <- '/project2/haky/temi/projects/l-TFPred/experiments/AR_phenotypes/data/processed_sumstats/'
out_folder <- asthma_folder
#gwas_folder <- '/project2/haky/temi/projects/l-TFPred/experiments/AR_phenotypes/data/processed_sumstats/'
run_script <- '/project2/haky/temi/projects/TFPred/scripts/clump_gwas_summary_statistics.sbatch'

out <- lapply(seq_along(asthma_phenotypes), function(i){
    pheno_folder <- glue('{out_folder}/{asthma_phenotypes[i]}')
    if(!dir.exists(pheno_folder)){
        dir.create(pheno_folder)
    }

    if(!file.exists(file.path(pheno_folder, glue('{asthma_phenotypes[i]}.gwas_sumstats.processed.txt')))){
        data.table::fread(file.path(out_folder, glue('{asthma_phenotypes[i]}.logistic.assoc.tsv.gz'))) %>%
            dplyr::mutate(SNP=paste(chr, pos, alt, ref, sep='_'), P=pval) %>%
            dplyr::rename(chrom=chr) %>%
            data.table::fwrite(., file.path(pheno_folder, glue('{asthma_phenotypes[i]}.gwas_sumstats.processed.txt')), col.names=T, row.names=F, sep='\t', quote=F)
    }
    
    print(glue('sbatch --export=phenotype={asthma_phenotypes[i]},pheno_folder={pheno_folder} {run_script}'))
})
```

```{r}
print(glue('{run_script} '))
```

```{r}
Filter(Negate(is.null), out) %>%
    do.call('rbind', .) %>%
    data.table::fwrite('/project2/haky/temi/projects/TFXcan-snakemake/metadata/metadata_gwas.txt', col.names=T, row.names=F, quote=F)
```

```{r}
testosterone_gwas <- bph_gwas %>%
    dplyr::mutate(p_value_adj = p.adjust(p_value, method = 'bonferroni')) %>%
    dplyr::filter(p_value_adj < 0.05)
data.table::fwrite(testosterone_gwas, file.path(work_dir, 'testosterone.gwas.sumstats.txt.gz'), col.names=T, row.names=F, quote=F, compress='gzip')
testosterone_gwas |> dim()
```

```{r}
data.frame(info = 'testosterone', gwas_sumstat = 'testosterone.gwas.sumstats.txt.gz') %>%
    data.table::fwrite('/project2/haky/temi/projects/TFXcan-snakemake/metadata/metadata_gwas.txt', col.names=T, row.names=F, quote=F)
```


### Analyse Enpact predictions

```{r}
enpact_folder <- '/project2/haky/temi/projects/TFXcan-snakemake/data/enpact_predictions'
```
```{r}
meta <- data.table::fread('/project2/haky/temi/projects/TFXcan-snakemake/metadata/metadata_gwas.txt')
meta
```

```{r}
enpact_predictions <- lapply(meta$phenotype, function(each_phenotype){
    list.files(enpact_folder, pattern = glue(".*({each_phenotype})_\\d.*"), full.names = T) %>%
        data.table::fread(.)
})
names(enpact_predictions) <- meta$phenotype
```

```{r}
# just making sure that all models are accounted for
cnames <- lapply(enpact_predictions, colnames)
enpact_models <- Reduce(f = intersect, x=cnames)[-1]
```

```{r}
out_1 <- lapply(enpact_predictions, function(each_dt){

    out <- lapply(enpact_models, function(each_model){
        each_dt %>%
            dplyr::pull(each_model)
    }) 
    names(out) <- enpact_models
    return(out)


    # return(out)
    
    # apply(combn(names(.), m=2), 2, function(each_col){
    #     nn <- paste(each_col, collapse=':')
    #     tt <- t.test(ar_out[[each_col[1]]],ar_out[[each_col[2]]])
    #     rr <- rbind(statistic = tt$statistic, pvalue = tt$p.value) |> 
    #         as.data.frame()
    #     names(rr) <- nn
    #     return(rr)
    # })
})

names(out_1) <- names(enpact_predictions)
```

```{r}

out_2 <- apply(combn(names(out_1), m=2), 2, function(each_col){
    nn <- paste(each_col, collapse=':')

    jj <- lapply(enpact_models, function(each_model){
        tt <- t.test(out_1[[each_col[1]]][[each_model]], out_1[[each_col[2]]][[each_model]])
        rr <- rbind(pvalue = tt$p.value) |> #statistic = tt$statistic, 
            as.data.frame()
        return(rr)
    })
    jj <- do.call('cbind', jj)
    names(jj) <- enpact_models
    rownames(jj) <- nn
    return(jj)
})

out_3 <- do.call('rbind', out_2) %>%
    as.data.frame() %>%
    tibble::rownames_to_column('pair_test')

```

```{r}
out_4 <- out_3 %>%
    tidyr::pivot_longer(cols=!pair_test, names_to = "enpact_model", values_to = "p_value") %>%
    dplyr::mutate(neg_log_p_value = -log10(p_value))
```

```{r}
out_4 %>%
    ggplot(., aes(pair_test, enpact_model, fill= neg_log_p_value)) + 
    geom_tile() +
    scale_fill_gradient(low="white", high="red") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
#enpact_models <- c('locus', 'AR_Prostate', 'STAT1_Lung', 'REST_Endometrium', 'RUNX1_Blood', 'E2F1_Mammary-Gland')

out <- lapply(enpact_predictions, function(each_dt){
    each_dt %>%
        dplyr::select(all_of(enpact_models))
})
```


```{r}
ar_out <- lapply(out, function(each_dt){
    each_dt %>%
        dplyr::pull(AR_Prostate)
})
```

```{r}
out_ttest <- apply(combn(names(ar_out), m=2), 2, function(each_col){
    nn <- paste(each_col, collapse=':')
    tt <- t.test(ar_out[[each_col[1]]],ar_out[[each_col[2]]])
    rr <- rbind(statistic = tt$statistic, pvalue = tt$p.value) |> 
        as.data.frame()
    names(rr) <- nn
    return(rr)
})

do.call('cbind', out_ttest) |> t() %>%
    as.data.frame() %>%
    tibble::rownames_to_column('pair_test') %>%
    dplyr::arrange(pvalue)
```



```{r}
lapply(out, function(each_dt){
    each_dt %>%
        dplyr::pull(AR_Prostate) %>%
        summary()
})
```



```{r}
library(data.table)
library(magrittr)
library(dplyr)
library(glue)
```


```{r}
dt <- data.table::fread('/project2/haky/Data/1000G/metadata/samples_information.txt')
```


```{r}
popuse <- c('AFR', 'AMR', 'EAS', 'EUR', 'SAS')
outdir <- '/project2/haky/Data/1000G/metadata'
dt %>% 
    dplyr::select(`Sample name`, `Superpopulation code`) %>% 
    dplyr::filter(`Superpopulation code` %in% popuse) %>% 
    dplyr::group_by(`Superpopulslation code`) %>% 
    group_walk(~ data.table::fwrite(as.data.frame(cbind(.x$`Sample name`, .x$`Sample name`)), glue('{outdir}/{.y$`Superpopulation code`}.samples.txt'), quote=F, col.names=F, row.names=F, sep='\t'))
```



```{r}
dt <- data.table::fread('/project/haky/users/temi/projects/data/GWAS/29892016-GCST006085-EFO_0001663.h.tsv.gz')
```

```{r}
dim(dt)
```

```{r}
head(dt)
```

```{r}
sv <- dt %>%
    dplyr::select(chrom = hm_chrom, pos = hm_pos, ref = hm_other_allele, alt = hm_effect_allele, beta = hm_beta, pval = p_value, variant_id = hm_variant_id, rsid = hm_rsid, se = standard_error, zscore = beta/standard_error) %>%
    dplyr::filter(!is.na(chrom) | !is.na(zscore))

sv <- sv[complete.cases(sv), ]
```

```{r}
dim(sv)
```

```{r}
sv %>%
    data.table::fwrite('/project/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/prostate_cancer_risk.liftover.logistic.assoc.tsv.gz', sep = '\t', quote = F, col.names = T, row.names = F)
```




```{r}
dt %>%
    dplyr::select(chrom, variant_id, rsid, pos, zscore, ref= A2, alt=A1) %>%
    data.table::fwrite('/project/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/prostate_cancer_risk.liftover.logistic.assoc.tsv.gz', sep = '\t', quote = F, col.names = T, row.names = F)
```

```{r}
dt <- data.table::fread('/project/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/prostate_cancer_risk.liftover.logistic.assoc.tsv.gz')
```

```{r}
# how many GWAS loci
mt <- dt %>%
    dplyr::mutate(adjP = p.adjust(pval, 'bonferroni'))
mt
```

```{r}
sum(mt$adjP < 5e-8)
```

```{r}
jt <- dt %>%
    dplyr::filter(!chrom %in% c('X', 'Y')) %>%
    dplyr::rename(CHR = chrom) %>%
    dplyr::mutate(CHR = as.numeric(CHR)) %>%
    dplyr::filter(CHR %in% c(1:22))
```

```{r}
pdf('/project/haky/users/temi/projects/TFXcan-snakemake/pcr.manhattan.pdf')
qqman::manhattan(jt, p = 'pval', bp = 'pos', snp = 'rsid')
dev.off()
```


```{r}
LD_block <- data.table::fread('/project/haky/users/temi/projects/TFXcan-snakemake/metadata/hg38_fourier_ls-all.bed') %>%
    dplyr::rename(chrom=V1, start=V2, stop=V3) %>%
    dplyr::mutate(split=1:nrow(.))
```


```{r}
library(data.table) |> suppressPackageStartupMessages()
library(tidyverse) |> suppressPackageStartupMessages()
library(susieR) |> suppressPackageStartupMessages()
library(glue) |> suppressPackageStartupMessages()
```


```{r}
sres <- readRDS('/project/haky/users/temi/projects/TFXcan-snakemake/data/finemapping/prostate_cancer_risk/prostate_cancer_risk.chr8.susie.RDS')
```


```{r}
susie_plot(sres$chr8_127147191_127147192$fitted_rss, y="PIP")
```


```{r}
srun <- readRDS(glue('/project/haky/users/temi/projects/TFXcan-snakemake/misc/chr11.susie.RDS'))
```


```{r}
susie_plot(srun$chr11_68238358_69701362$fit, 'PIP')
```


```{r}
susie_plot(srun$chr11_101460390_104088908$fit, 'PIP')
```


```{r}
# chr11_68238358_69701362
susie_plot(srun$chr11_68238358_69701362$fit, 'PIP')
```


```{r}
srun$chr11_70855_1192365$fit
```


```{r}
dt <- data.table::fread('/project/haky/users/temi/projects/TFXcan-snakemake/data/PCR_GWAS_2024-06-04/predictdb/prostate_cancer_risk.enpact_scores.txt')

mt <- data.table::fread('/project/haky/users/temi/projects/TFXcan-snakemake/data/PCR_GWAS_2024-06-04/predictdb/prostate_cancer_risk.tf_tissue_annot.txt')
```

```{r}
mt %>%
    dplyr::mutate(gene_id = gsub('-', '', gene_id), gene_name = gsub('-', '', gene_name)) %>%
    data.table::fwrite('/project/haky/users/temi/projects/TFXcan-snakemake/data/PCR_GWAS_2024-06-04/predictdb/prostate_cancer_risk.FT.tf_tissue_annot.txt', col.names = T, row.names = T, sep = '\t', quote = F)
```


```{r}
dt %>% dplyr::mutate(NAME = gsub('-', '', NAME)) %>%
    data.table::fwrite('/project/haky/users/temi/projects/TFXcan-snakemake/data/PCR_GWAS_2024-06-04/predictdb/prostate_cancer_risk.FT.enpact_scores.txt', col.names = T, row.names = F, sep = '\t', quote = F)
```

## Standing height GWAS

```{r}
# chrom   pos     alt     ref     variant_id      beta    se      pval    zscore
height_ss <- data.table::fread('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/standing_height.gwas_sumstats.tsv.gz') %>%
    dplyr::rename(chrom = chromosome, pos = base_pair_location, ref = other_allele, alt = effect_allele, beta = beta, se = standard_error) %>% 
    dplyr::filter(nchar(alt) == 1 & nchar(ref) == 1) %>%
    dplyr::mutate(zscore = beta/se, variant_id = paste(chrom, pos, ref, alt, sep = '_'), pval = 10^-(neg_log_10_p_value))
```


```{r}
height_ss %>% 
    data.table::fwrite('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/standing_height.gwas_sumstats.processed.txt.gz', col.names = T, row.names = F, sep = '\t', quote = F, compress = 'gzip')
```

```{r}
xdt <- data.frame(num= 1:26, let = LETTERS, let2 = letters) %>%
    dplyr::mutate(vv = paste(let, let2, num, sep = '_'))
xdt
```


```{r}
dt_height 
```

# CXCL9 protein levels

https://www.ebi.ac.uk/gwas/studies/GCST90428427

Download the harmonized GWAS file. 

```{r}
# chrom   pos     alt     ref     variant_id      beta    se      pval    zscore
CXCL9_protein_levels <- data.table::fread('/beagle3/haky/users/temi/projects/TFXcan-snakemake/minimal/data/GCST90428427.h.tsv.gz') %>%
    dplyr::rename(chrom = chromosome, pos = base_pair_location, ref = other_allele, alt = effect_allele, beta = beta, se = standard_error, pval = p_value) %>% 
    dplyr::filter(nchar(alt) == 1 & nchar(ref) == 1) %>%
    dplyr::mutate(zscore = beta/se, variant_id = paste(chrom, pos, ref, alt, sep = '_'))
```


```{r}
CXCL9_protein_levels %>% 
    data.table::fwrite('/beagle3/haky/users/temi/projects/TFXcan-snakemake/minimal/data/CXCL9_protein_levels.gwas_sumstats.processed.hg38.txt.gz', col.names = T, row.names = F, sep = '\t', quote = F, compress = 'gzip')
```


## Type 2 diabetes

Downloaded harmonized data from https://www.ebi.ac.uk/gwas/studies/GCST90296697

Paper: https://pubmed.ncbi.nlm.nih.gov/38182742/

```{bash}
wget https://ftp.ebi.ac.uk/pub/databases/gwas/summary_statistics/GCST90296001-GCST90297000/GCST90296697/harmonised/GCST90296697.h.tsv.gz -O /beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.males.gwas_sumstats.tsv.gz

wget https://ftp.ebi.ac.uk/pub/databases/gwas/summary_statistics/GCST90296001-GCST90297000/GCST90296698/harmonised/GCST90296698.h.tsv.gz -O /beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.females.gwas_sumstats.tsv.gz
```

```{r}
# chrom   pos     alt     ref     variant_id      beta    se      pval    zscore
t2d_males_ss <- data.table::fread('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.males.gwas_sumstats.tsv.gz') %>%
    dplyr::rename(chrom = chromosome, pos = base_pair_location, ref = other_allele, alt = effect_allele, beta = beta, se = standard_error, pval = p_value) %>% 
    dplyr::filter(nchar(alt) == 1 & nchar(ref) == 1) %>%
    dplyr::mutate(zscore = beta/se, variant_id = paste(chrom, pos, ref, alt, sep = '_'))
```




```{r}
t2d_males_ss %>% 
    data.table::fwrite('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.males.gwas_sumstats.processed.txt.gz', col.names = T, row.names = F, sep = '\t', quote = F, compress = 'gzip')
```

```{r}
# chrom   pos     alt     ref     variant_id      beta    se      pval    zscore
t2d_females_ss <- data.table::fread('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.females.gwas_sumstats.tsv.gz') %>%
    dplyr::rename(chrom = chromosome, pos = base_pair_location, ref = other_allele, alt = effect_allele, beta = beta, se = standard_error, pval = p_value) %>% 
    dplyr::filter(nchar(alt) == 1 & nchar(ref) == 1) %>%
    dplyr::mutate(zscore = beta/se, variant_id = paste(chrom, pos, ref, alt, sep = '_'))
```


```{r}
t2d_females_ss %>% 
    data.table::fwrite('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.females.gwas_sumstats.processed.txt.gz', col.names = T, row.names = F, sep = '\t', quote = F, compress = 'gzip')
```

```{r}
hist(t2d_females_ss$pval)
```

```{r}
max(t2d_females_ss$pval)
```

## T2D suzuki


```{r}
# chrom   pos     alt     ref     variant_id      beta    se      pval    zscore
t2d_suzuki_ss <- data.table::fread('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.suzuki.gwas_sumstats.tsv.gz') %>%
    dplyr::rename(chrom = Chromsome, pos = Position, ref = NonEffectAllele, alt = EffectAllele, beta = Beta, se = SE, pval = Pval) %>% 
    dplyr::filter(nchar(alt) == 1 & nchar(ref) == 1) %>%
    dplyr::mutate(zscore = beta/se, variant_id = paste(chrom, pos, ref, alt, sep = '_'))
```

```{r}
# library(bigsnpr)
```

```{r}

# bx <- t2d_suzuki_ss[1:100, ] %>% dplyr::mutate(chr = paste('chr', chr, sep = ''))
# xx <- snp_modifyBuild(
#   bx,
#   '/project2/haky/Data/liftover/liftOver2',
#   from = "hg19",
#   to = "hg38",
#   check_reverse = TRUE,
#   local_chain = '/project2/haky/Data/liftover/ucsc_chainfiles/hg19ToHg38.over.chain.gz',
#   #base_url = "https://hgdownload.soe.ucsc.edu/goldenPath/"
# )
```


```{r}
t2d_suzuki_ss %>% 
    data.table::fwrite('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.suzuki.gwas_sumstats.processed.txt.gz', col.names = T, row.names = F, sep = '\t', quote = F, compress = 'gzip')
```

```{r}
# liftover

chainObject <- rtracklayer::import.chain('/project2/haky/Data/liftover/ucsc_chainfiles/hg19ToHg38.over.chain')

```

```{r}
grObject <- GenomicRanges::GRanges(seqnames = t2d_suzuki_ss$chrom, ranges = IRanges(start = t2d_suzuki_ss$pos, end = t2d_suzuki_ss$pos))

mcols(grObject) <- t2d_suzuki_ss[, c('variant_id', 'ref', 'alt', 'zscore', 'pval', 'se', 'beta', 'pos', 'chrom')]

seqlevelsStyle(grObject) <- "UCSC"
results <- as.data.frame(rtracklayer::liftOver(grObject, chainObject))
```

```{r}
head(results)
```

```{r}
results %>%
    dplyr::mutate(variant_id = paste(chrom, start, ref, alt, sep = '_'), pos = start) %>%
    dplyr::select(chrom, pos, variant_id, ref = ref, alt = alt, zscore = zscore, pval = pval, se = se, beta = beta) %>% 
    data.table::fwrite('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.suzuki.gwas_sumstats.hg38.processed.txt.gz', col.names = T, row.names = F, sep = '\t', quote = F, compress = 'gzip')
```

```{r}
data.table::fread('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.suzuki.gwas_sumstats.hg38.processed.txt.gz') %>%
    dplyr::filter(chrom %in% c(1:3)) %>%
    data.table::fwrite('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/type_2_diabetes.suzuki.mini.gwas_sumstats.hg38.processed.txt.gz', col.names = T, row.names = F, sep = '\t', quote = F, compress = 'gzip')

```



```{r}
rt <- data.table::fread('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/T2D_2025-01-24/collection/t2d_suzuki.filteredGWAS.topSNPs.txt.gz')

et <- data.table::fread('/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/T2D_2025-01-24/collection/t2d_suzuki.EnformerLoci.topSNPs.txt', header = F)
```


```{r}
rt %>% dplyr::arrange(pval) %>%
        dplyr::slice(1:100)
```

## Pancreatic beta cell function

Source:
Here: wget https://api.kpndataregistry.org/api/d/ChTEgN -O pancreatic_beta_cell_function.zip

Then: unzip pancreatic_beta_cell_function.zip -d pancreatic_beta_cell_function

Details:

1. Genome build is hg37. So, you need to liftover to hg38.
2. There are 8 metabolic traits here. I have to process them individually.

```{r}
pbcf_directory <- '/beagle3/haky/users/temi/projects/TFXcan-snakemake/data/sumstats/pancreatic_beta_cell_function'
pbcf_files <- list.files(pbcf_directory, pattern = '.*', full.names = T)

# I need the names from the file names 
pbcf_names <- basename(pbcf_files) %>%
    gsub('.filthetmafn.rsid.selectedcolumns', '', .)
names(pbcf_files) <- pbcf_names
```
```{r}
# chrom   pos     alt     ref     variant_id      beta    se      pval    zscore
pbcf_sumstats <- purrr::map(names(pbcf_files), function(nm){
    dt <- data.table::fread(pbcf_files[nm])
    # using tidyr::separate was too slow
    sp <- stringi::stri_split_fixed(dt$MarkerName, pattern = ':') %>% 
        do.call('rbind', .) %>% as.data.table()
    colnames(sp) <- c('chrom', 'pos')

    dt %>%
        dplyr::rename(alt = Allele1, ref = Allele2, beta = Effect, se = StdErr, pval = `P-value`) %>% 
        dplyr::filter(nchar(alt) == 1 & nchar(ref) == 1) %>%
        dplyr::bind_cols(sp) %>%
        dplyr::mutate(alt = toupper(alt), ref = toupper(ref), zscore = beta/se, variant_id = paste(chrom, pos, ref, alt, sep = '_'), pos = as.numeric(pos))
}, .progress = TRUE)

names(pbcf_sumstats) <- pbcf_names
```


```{r}
# liftover
chainObject <- rtracklayer::import.chain('/project2/haky/Data/liftover/ucsc_chainfiles/hg19ToHg38.over.chain')
```

```{r}

purrr::map(names(pbcf_sumstats), function(nm){
    pbcf_ss <- pbcf_sumstats[[nm]]
    grObject <- GenomicRanges::GRanges(seqnames = pbcf_ss$chrom, ranges = IRanges(start = pbcf_ss$pos, end = pbcf_ss$pos))

    mcols(grObject) <- pbcf_ss[, c('variant_id', 'ref', 'alt', 'zscore', 'pval', 'se', 'beta', 'pos', 'chrom')]

    ensembldb::seqlevelsStyle(grObject) <- "UCSC"
    results <- as.data.frame(rtracklayer::liftOver(grObject, chainObject))
    
    results %>%
        dplyr::mutate(variant_id = paste(chrom, start, ref, alt, sep = '_'), pos = start) %>%
        dplyr::select(chrom, pos, variant_id, ref = ref, alt = alt, zscore = zscore, pval = pval, se = se, beta = beta) %>% 
        data.table::fwrite(glue('{pbcf_directory}/{nm}.gwas_sumstats.hg38.processed.txt.gz'), col.names = T, row.names = F, sep = '\t', quote = F, compress = 'gzip')
}, .progress = TRUE)

```

```{r}
# how many GWAS loci per LD block
ld_blocks <- data.table::fread('/beagle3/haky/users/temi/projects/TFXcan-snakemake/metadata/EUR_LDBlocks.hg38.updated.bed') %>% setNames(c('ld_chrom', 'ld_start', 'ld_stop'))

ld_splits <- ld_blocks %>%
    dplyr::group_split(ld_chrom, ld_start, ld_stop) 
```

```{r}
# how many top GWAS loci per LD block
pbcf_loci <- purrr::map(names(pbcf_sumstats), function(nm){
    pbcf_ss <- pbcf_sumstats[[nm]] %>%
        dplyr::filter(pval < 5e-8) %>%
        dplyr::mutate(chrom = paste('chr', chrom, sep = ''))

    xd_splits <- lapply(ld_splits, function(each_split){
        xd <- dplyr::filter(pbcf_ss, chrom == each_split$ld_chrom[1], pos >= each_split$ld_start[1] & pos <= each_split$ld_stop[1])
        if(nrow(xd) == 0){
            return(NULL)
        } else {
            return(xd)
        }
    })

    xd_splits <- Filter(Negate(is.null), xd_splits)
}, .progress = TRUE)
```

```{r}
sapply(pbcf_loci, length) |> sum()
```




```{r}
chainfile <- rtracklayer::import.chain('/project/haky/data/liftover/chainfiles/ucsc/hg19ToHg38.over.chain')
```

```{r}
# bc <- data.table::fread("/project/haky/users/temi/projects/data/sumstats/icogs_onco_gwas_meta_overall_breast_cancer_summary_level_statistics.txt")
```

```{bash}
awk '{print $9, $10, $2, $3, $1, $5, $6, $7}' /project/haky/users/temi/projects/data/sumstats/icogs_onco_gwas_meta_overall_breast_cancer_summary_level_statistics.txt > breast_cancer.sumstats.hg19.txt;
gzip -c breast_cancer.sumstats.hg19.txt > breast_cancer.sumstats.hg19.txt.gz
```

```{bash}
40 41 42 43 44 45 46
```

```{bash}
awk '{print $9, $10, $2, $3, $1, $5, $6, $7, $40, $41, $42, $43, $44, $45, $46}' /project/haky/users/temi/projects/data/sumstats/icogs_onco_gwas_meta_overall_breast_cancer_summary_level_statistics.txt | head
```

```{bash}

# $2, $3, $5, $6, $7
awk '{print $9, $10, $1, $40, $41, $42, $43, $44, $45, $46}' /project/haky/users/temi/projects/data/sumstats/icogs_onco_gwas_meta_overall_breast_cancer_summary_level_statistics.txt > breast_cancer.sumstats.hg19.txt;
gzip -c breast_cancer.sumstats.hg19.txt > breast_cancer.sumstats.hg19.txt.gz
```


```{r}
bc <- data.table::fread("/project/haky/users/temi/projects/data/sumstats/breast_cancer.sumstats.hg19.txt.gz")
```


```{r}
# bc2 <- bc %>%
#     stats::setNames(nm = c('chrom', 'pos', 'alt', 'ref', 'variant_id', 'beta', 'se', 'pval')) %>%
#     dplyr::mutate(zscore = beta/se, chr = paste('chr', chrom, sep = '')) %>%
#     dplyr::filter(chrom %in% 1:22, nchar(ref) == 1, nchar(alt) == 1)


bc2 <- bc %>%
    stats::setNames(nm = c('chrom', 'pos', 'variant_id', 'alt', 'ref', 'beta', 'var', 'se', 'chi', 'pval')) %>%
    dplyr::mutate(zscore = sqrt(chi), chr = paste('chr', chrom, sep = '')) %>%
    dplyr::select(-c(chi, var)) %>%
    dplyr::filter(chrom %in% 1:22, nchar(ref) == 1, nchar(alt) == 1)
```

```{r}
data.table::fwrite(bc2, "/project/haky/users/temi/projects/data/sumstats/breast_cancer.sumstats.hg19.renamed.txt.gz", sep = '\t', compress = 'gzip', row.names = F, col.names = T)
```

```{r}
bc2_granges <- with(bc2[1:20, ], GenomicRanges::GRanges(seqnames = chr, IRanges(pos, pos + 1), strand = '*', chrom, pos, alt, ref, variant_id, beta, se, pval, zscore))
bc2_granges
```


```{r}
bc2_liftover <- rtracklayer::liftOver(bc2_granges, chainfile)
bc2_liftover
```


```{r}
lf <- as.data.frame(bc2_liftover) %>%
    dplyr::select(-c(group, group_name, width, strand, pos, seqnames)) %>%
    dplyr::select(chrom, start, end, alt, ref, variant_id, beta, se, pval, zscore)

lf
```


```{r}
grObject <- GRanges(seqnames=c("chr1"), ranges=IRanges(start=226061851, end=226071523))
results <- as.data.frame(liftOver(grObject, chainfile))
```


```{bash}
sbatch /project/haky/users/temi/projects/TFXcan-snakemake/workflow/helpers/liftover.sbatch /project/haky/users/temi/projects/data/sumstats/breast_cancer.sumstats.hg19.renamed.txt.gz /project/haky/data/liftover/chainfiles/ucsc/hg19ToHg38.over.chain /project/haky/users/temi/projects/data/sumstats/breast_cancer.sumstats.hg38.renamed.txt.gz
```


```{r}
bc3 <- data.table::fread("/project/haky/users/temi/projects/data/sumstats/breast_cancer.sumstats.hg38.renamed.txt.gz") 
```


```{r}
bc2 %>% dplyr::filter(chrom == 9 & between(pos, 16912915, 16912925))
```

```{r}
bc2 %>% dplyr::filter(chrom == 9 & between(pos, 16914710, 16914720))
```



```{r}
bc3 %>% dplyr::filter(chrom == 9 & between(pos, 16912915, 16912925))
```

```{r}
bc3 %>% dplyr::filter(chrom == 9 & between(pos, 16914710, 16914720))
```


```{r}
bc3 %>% dplyr::filter(chrom == 9) %>% head()
```


```{bash}
zgrep -E chr9_16912921_A_G chr9.sumstats.txt.gz 
```


```{bash}
zgrep 17_41961940_ chr17.sumstats.txt.gz 
```


```{r}
mt <- data.table::fread('/project/haky/users/temi/projects/TFXcan-snakemake/metadata/models.enpact.185.txt') %>%
    dplyr::distinct()
```


```{r}
mt %>% data.table::fwrite('/project/haky/users/temi/projects/TFXcan-snakemake/metadata/models.enpact.185.txt', row.names = F, col.names = T, sep = '\t', quote = F)
```


```{r}
jt <- data.table::fread('/project/haky/users/temi/projects/TFXcan-snakemake/data/BC_GWAS/enpact_predictions/breast_cancer/NA20832.breast_cancer.aggByCollect.2024-06-06.csv.gz')
```


### create fictional filtered GWAs file
```{r}
# bc3 %>% dplyr::filter(
#     chrom == 9 & pos == 16912921
#     chrom == 9 & pos == 16914718,
#     chrom == 17 & pos == 41961970,
#     chrom == 17 & pos == 43582425
#     )

ft <- dplyr::bind_rows(
    bc3 %>% dplyr::filter(chrom == 9 & pos == 16912921),
bc3 %>% dplyr::filter(chrom == 9 & pos == 16914718),
bc3 %>% dplyr::filter(chrom == 17 & pos == 41961940),
bc3 %>% dplyr::filter(chrom == 17 & pos == 43582425),
)

ft
```

```{r}
ft %>%
    dplyr::rename(a0 = ref, a1 = alt) %>%
    dplyr::mutate(rsid = paste(chrom, pos, a0, a1, sep = ':'), adjP = pval, gwas_significant = ifelse(pval < 5e-8, 'YES', "NO"), varID = paste(chrom, pos, a0, a1, sep = '_')) %>%
    dplyr::select(-variant_id) %>%
    data.table::fwrite('/project/haky/users/temi/projects/TFXcan-snakemake/data/BC_GWAS/collection/breast_cancer/breast_cancer.filteredGWAS.txt.gz', sep = '\t', quote = F, row.names = F, col.names = T, compress = 'gzip')
```


```{r}
bc3 %>% dplyr::filter(
    chrom == 9 & pos == 16914718
    )
```



```{r}
library(susieR)
```


```{r}
srun <- readRDS('/project/haky/users/temi/projects/TFXcan-snakemake/data/BC_GWAS/finemapping/breast_cancer/breast_cancer.chr4.susie.RDS')

sfit <- srun$`chr4_37253425_38905401`$fit
```


```{r}
susie_plot(sfit, 'PIP')
```


```{r}
summary(sfit)
```